Technical Report for Real-Time Certified Probabilistic Pedestrian Forecasting

### 2017-06-17 final/technicalreportarxiv/pedestrian_forecasting_iros_2017-technicalreport.tex
Revision with Citation.

% numbers option provides compact numerical references in the text. 
%\usepackage[numbers]{natbib}
%\usepackage{multicol}
\usepackage[bookmarks=true]{hyperref}
\newcommand{\Ram}[1]{\textcolor{red}{#1}}
\newcommand{\Hoj}[1]{\textcolor{blue}{#1}}
\newcommand{\rtext}[1]{\textcolor{red}{#1}}

@@

% paper title
\title{\textbf{Technical Report:} \\ Real-Time Certified Probabilistic Pedestrian Forecasting}

%%% first author
\author{Henry O. Jacobs, Owen K. Hughes, Matthew Johnson-Roberson, and Ram Vasudevan
 \thanks{ H.O. Jacobs, O. Hughes, M. Johnson-Roberson, and R.~Vasudevan are with the University of Michigan, Ann Arbor, MI 48109
{\scriptsize \{\texttt{hojacobs,owhughes,mattjr,ramv}\}~\texttt{@umich.edu}}}
\thanks{This work was supported by Ford Motor Company.}
}

@@

%The approach presented in \cite{Kitani2012} has been generalized in a variety of ways.
This approach has been generalized in a variety of ways.
For example, time-dependent information, such as traffic signals, are incorporated in \cite{Karasev2016}, by relaxing the Markov property and considering a switched Markov process.
Other generalizations include replacing the finite-state space with a continuous one, and using a Markov jump process in the motion model.
Unfortunately the desired posteriors are difficult to compute in closed form, and as a result a sampling based method.
The resulting accuracy of such methods, which can come at a large computational expense, can only be known in a probabilistic sense in that the error bounds are themselves random variables.
%Moreover, accuracy can come at a large computational expense.
% which is prohibitive during real-time control.

One limitation of IOC models occurs in cases where there are locally optimal solutions between a start and end goals which can yield non-robust and imprecise behavior.
This can occur when agents make sharp turns due to intermediate criteria on the way toward reaching their final destination.
To address this, \cite{Ballan2016} adopt an empiricists approach, computing ``turning maps'' and attempting to infer how agents behave in a given patch.
The motion model is a Markov jump process and the relevant posteriors are approximated using sample based techniques similar to \cite{Karasev2016}.
The objective of \cite{Ballan2016} is not only prediction, but the development of a motion model learned on one scene that could then subsequently be transferred to other scenes. 
This requires representations of ``objects'' in the scene that do not depend rigidly on the finitely many labels an engineer managed to think of in a late-night brainstorming session.
%For example, such an algorithm should be able to infer how to behave near a roundabout by using prior knowledge of more fundamental building blocks like pavement, curbs, and asphalt.

@@

Many have approached pedestrian forecasting by deriving their motion model from interactions between pedestrians using physically motivated methods \cite{Helbing1995,Xu2012}. 
	Several models derive their motion models from \cite{Helbing1995} by incorporating collision avoidance through an interaction potential \cite{Pellegrini2009,Yamaguchi2011,Yi2016}. 
	However this method suffers from not planning for other pedestrian positions at future times. 
	Others take optical flow as input \cite{Hospedales2009,Wang2009,Emonet2011}.
	These approaches use variants of Hierarchical Dirichlet Processes on discretized optical flow to determine temporal motifs (i.e. classes of motion within the scene), or on a Markov model. 
	 These models are not agent based, and the lack of an explicit motion model limits their predictive power. 
	 Recently methods have been developed to predict trajectories by introducing and sampling Anticipatory Temporal Conditional Random Fields which incorporate learned affordances based on observed objectives within the scene \cite{Koppula2016}. 
	 Others create agent-based models based on Gaussian Processes, though they suffer issues when trained on discretized trajectories \cite{Tay2008,Wang2008,Trautman2015}. 
	Most recently a method using Long Short-Term Memory (lstm) to learn pedestrian motion models without making assumptions about the manner in which agents interact is the state-of-the-art algorithm for socially based models and has a rapid computation time \cite{Alahi2016}.

@@

  \subsection{Learning the Vector Fields}
 We begin by identifying the number of possible vector-fields.
  %In principal, any clustering algorithm could accomplish this.
To do, this we use a clustering algorithm on the trajectories using Affinity Propagation \cite{FreyDueck2007} and a custom distance measure defined on the endpoints of trajectories. 
	Let one trajectory $A$ start at $(x_{A,\mathrm{start}}, y_{A,\mathrm{start}})$ and end at  $(x_{A,\mathrm{end}}, y_{A,\mathrm{end}})$, and another trajectory $B$ start at $(x_{B,\mathrm{start}}, y_{B,\mathrm{start}})$ and end at $(x_{B,\mathrm{end}}, y_{B,\mathrm{end}})$ . 
	We define the points $\mathbf{a}_1 = (x_{A, \mathrm{start}}, y_{A, \mathrm{start}}, x_{A, \mathrm{end}}, y_{A, \mathrm{end}})$, $\mathbf{a}_2 = (x_{A, \mathrm{end}}, y_{A, \mathrm{end}}, x_{A, \mathrm{start}}, y_{A, \mathrm{start}})$, and $\mathbf{b} = (x_{B, \mathrm{start}}, y_{B, \mathrm{start}}, x_{B, \mathrm{end}}, y_{B, \mathrm{end}})$. 
	We then define our distance measure as $d(A, B) :=  \min \left\{ d_e(\mathbf{a}_1, \mathbf{b}), d_e(\mathbf{a}_2, \mathbf{b}) \right\}$, for the euclidean distance $d_e$ in $\mathbb{R}^4$. 
	This function measures the distance between the endpoints irrespective of their ordering, which means that a trajectory that starts from point A and ends at point B will be close to a trajectory that starts from point B and ends at point A.
	The scale of the datasets we tested on had large enough spatial scale that clustering based on endpoints captured people moving from destination to destination, e.g. from a storefront to the sidewalk at the edge of a scene. On this dataset, distance functions that utilize the entire trajectory such as DTW, LCSS, and PF from \cite{Morris2009} and TRACLUS from \cite{Lee2007} did not identify pedestrian intent as well as our method did. While it is unclear why TRACLUS underperformed on our data, PF, DTW, and LCSS appeared to put trajectories that were very similar for periods of time together even though they had very different intents.
	This clustering of the end-points induces a clustering of the trajectories.
  Suppose we obtain clusters $S_1, \dots, S_n$ consisting of trajectories from our data set, as well as a set of unclassified trajectories, $S_0$.

@@

Note that the implementation of the algorithm in \cite{Kitani2012} required the endpoint of each test trajectory. 
Without this information the implementation of the predictor provided by the authors devolved into a random walk.
Neither our algorithm nor the random walk were provided this information.

The output distributions of the four algorithms were compared using their integrals over the cells of a regular grid over our domain. 
These integrals are used to visualize the distributions in Figures \ref{fig:bookstore-1-2} and \ref{fig:death-1-2}.
%In keeping with our measurement model, the ground truth compared against was a grid that was identically zero except on a bounding box around the measurement at time $t$. 
%The ground truth bounding box at each time instance were set equal to the average bounding box dimension centered at the 
%were the average bounding box dimensions in the training data set. 
Because our predictions all share the same scale, we amalgamate all of the prediction and truth values for all of the simulated agents at a given time step and generate ROC curves.
In our analysis, we sought a metric that evaluated the utility of prediction algorithms in the autonomous vehicle context. 
	In this instance it is critical that the generated set of predictions contains the ground-truth observed trajectory while including as few false positive detections as possible.
	ROC curves which plot the true positive rate (i.e. the proportion of the ground truth trajectory that is correctly identified) against the false positive rate (i.e. the proportion of the generated set that does not correspond to the ground truth trajectory), evaluate this aforementioned safety criteria for autonomous vehicles exactly. 
	The Area Under the Curve (AUC) is a standard measure of the quality of a predictor.  
	The closer that this AUC is to one, the better the prediction algorithm.
	Figure \ref{fig:auc_vs_time} shows the analysis of the AUC of each algorithm versus time.

In addition, we evaluated the Modified Hausdorff Distance (MHD) from the ground truth trajectory to a sample from the predictions at each time to provide a geometric measure of how accurate the predictions are. 
	Figure \ref{fig:mhd_vs_time} shows MHD plotted against time.
	Though popular in evaluating the geometric proximity of the ground truth trajectory and a predicted set, it is not the most appropriate way to evaluate the utility of an algorithm in the autonomous vehicle context. 
	Specifically consider the instance of a generated set of predictions which does not contain the ground-truth observed trajectory but is close to the ground-truth trajectory geometrically. 
	If this generated set of predictions was used in the autonomous vehicle context, it would not preclude certain portions of the space where the person was occupying. 
	Notice that the true positive rate would be zero meaning that the generated set of predictions was not useful.
	Whereas the MHD would describe this generated set of predictions as informative.
Our predictor behaves better than any of the other compared methods at moderately large $t$.
%, where we capture the behavior of the agent without suffering the misplaced confidence observed in Kitani et al.'s predictions. 
%Their encoding of velocity results in the prediction lagging behind the actual observed motion. 
This is despite providing the algorithm in \cite{Kitani2012} with the specific final location of each agent.
%Even given significantly better information, Kitani et al. generates lower quality predictions than our algorithm at all time scales.

@@

\begin{figure}
	\centering
	\includegraphics[width=0.9\linewidth]{./figures/the_results.eps}
	\caption{A comparison of the AUC of the various algorithms. Note that the initial dip in the performance of \cite{Kitani2012} is due to their confidence in their initial estimate. We sampled the S-LSTM \cite{Alahi2016} model $100$ times to extract a less concentrated probability distribution from their algorithm.}
%	\normalmarginpar\marginnote{\#4.7}
	\label{fig:auc_vs_time}
	\vspace*{-0.75cm}
\end{figure}

\begin{figure}
	\centering
	\includegraphics[width=0.9\linewidth]{./figures/MHD_results.eps}
	\caption{A comparison of the MHD from the ground truth of the pedestrian to a 1000 point samples from each distribution. The method from \cite{Alahi2016} does well at short time scales since it places a highly confident distribution at the given initial position of the pedestrian, but the method developed in this paper outperforms all others at intermediate times. \cite{Kitani2012}, which requires the end point of each trajectory, outperforms all other algorithms at longer time scales since they assume that the end point of the trajectory is known.}
%	\reversemarginpar\marginnote{\#4.8}
	\label{fig:mhd_vs_time}
	\vspace*{-0.7cm}
\end{figure}

@@

\section{Conclusion} 
\label{sec:conclusion}
This paper presents a real-time approach to probabilistic pedestrian modeling. 
We demonstrate the ability to accurately predict the final location of a pedestrian with superior accuracy compared to a state-of-the-art with additional contributions in two critical areas: 1) prediction speed is efficient to enable real-time control and 2) our error bound on the predictions accurately reflects potential paths for the human. 
These two features have great utility for robotic applications, and we see the integration of such techniques with autonomous system control as one pathway to enable safe operation in human-robot spaces.

>>>>

% numbers option provides compact numerical references in the text. 
%\usepackage[numbers]{natbib}
%\usepackage{multicol}
\usepackage{hyperref}

@@

% paper title
\title{Technical Report for Real-Time Certified Probabilistic Pedestrian Forecasting}

%%% first author
\author{Henry O. Jacobs, Owen K. Hughes, Matthew Johnson-Roberson, and Ram Vasudevan
 \thanks{ H.O. Jacobs, O. Hughes, M. Johnson-Roberson, and R.~Vasudevan are with the University of Michigan, Ann Arbor, MI 48109
{\scriptsize \{\texttt{hojacobs,owhughes,mattjr,ramv}\}~\texttt{@umich.edu}}}
\thanks{This material is based upon work supported by Ford Motor Company and the National Science Foundation under Grant No. 1562612.}
}

@@

%The approach presented in \cite{Kitani2012} has been generalized in a variety of ways.
This approach has been generalized in a variety of ways.
For example, time-dependent information, such as traffic signals, are incorporated in \cite{Karasev2016}, by relaxing the Markov property and considering a switched Markov process.
Other generalizations include replacing the finite-state space with a continuous one, and using a Markov jump process in the motion model.
Unfortunately the desired posteriors are difficult to compute in closed form, and as a result use sampling based methods.
The resulting accuracy of such methods, which can come at a large computational expense, can only be known in a probabilistic sense in that the error bounds are themselves random variables.
%Moreover, accuracy can come at a large computational expense.
% which is prohibitive during real-time control.

A limitation of IOC models occurs when there are locally optimal solutions between a start and end goals that can yield non-robust and imprecise behavior.
This occurs when agents make sharp turns due to intermediate criteria on the way toward reaching their final destination.
To address this, \cite{Ballan2016} adopt an empiricists approach, computing ``turning maps'' and attempting to infer how agents behave in a given patch.
The motion model is a Markov jump process and the relevant posteriors are approximated using sample based techniques similar to \cite{Karasev2016}.
The objective of \cite{Ballan2016} is not only prediction, but the development of a motion model learned on one scene that could then subsequently be transferred to other scenes. 
This requires representations of ``objects'' in the scene that do not depend rigidly on the finitely many labels an engineer managed to think of in a late-night brainstorming session.
%For example, such an algorithm should be able to infer how to behave near a roundabout by using prior knowledge of more fundamental building blocks like pavement, curbs, and asphalt.

@@

Many have approached pedestrian forecasting by deriving their motion model from interactions between pedestrians using physically motivated methods \cite{Helbing1995,Xu2012}. 
	Several models derive their motion models from \cite{Helbing1995} by incorporating collision avoidance through an interaction potential \cite{Pellegrini2009,Yamaguchi2011,Yi2016}. 
	However this method suffers from not planning for other pedestrian positions at future times. 
	Others take optical flow as input \cite{Hospedales2009,Wang2009,Emonet2011}.
	These approaches use variants of Hierarchical Dirichlet Processes on discretized optical flow to determine temporal motifs (i.e. classes of motion within the scene), or on a Markov model. 
	 These models are not agent based, and the lack of an explicit motion model limits their predictive power. 
	Recently, methods have been developed to predict trajectories by introducing and sampling Anticipatory Temporal Conditional Random Fields which incorporate learned affordances based on observed objectives within the scene \cite{Koppula2016}.
	 Others create agent-based models based on Gaussian Processes which perform poorly when trained on discretized trajectories \cite{Tay2008,Wang2008,Trautman2015}.
	Most recently, a method using Long Short-Term Memory (LSTM) was proposed to learn pedestrian motion models without making assumptions about the manner in which agents interact while having a rapid computation time \cite{Alahi2016}.

@@

  \subsection{Learning the Vector Fields}
 We begin by identifying the number of possible vector-fields.
  %In principal, any clustering algorithm could accomplish this.
To do, this we use a clustering algorithm on the trajectories using Affinity Propagation \cite{FreyDueck2007} and a custom distance measure defined on the endpoints of trajectories. 
	Let one trajectory $A$ start at $(x_{A,\mathrm{start}}, y_{A,\mathrm{start}})$ and end at  $(x_{A,\mathrm{end}}, y_{A,\mathrm{end}})$, and another trajectory $B$ start at $(x_{B,\mathrm{start}}, y_{B,\mathrm{start}})$ and end at $(x_{B,\mathrm{end}}, y_{B,\mathrm{end}})$ . 
	We define the points $\mathbf{a}_1 = (x_{A, \mathrm{start}}, y_{A, \mathrm{start}}, x_{A, \mathrm{end}}, y_{A, \mathrm{end}})$, $\mathbf{a}_2 = (x_{A, \mathrm{end}}, y_{A, \mathrm{end}}, x_{A, \mathrm{start}}, y_{A, \mathrm{start}})$, and $\mathbf{b} = (x_{B, \mathrm{start}}, y_{B, \mathrm{start}}, x_{B, \mathrm{end}}, y_{B, \mathrm{end}})$. 
	We then define our distance measure as $d(A, B) :=  \min \left\{ d_e(\mathbf{a}_1, \mathbf{b}), d_e(\mathbf{a}_2, \mathbf{b}) \right\}$, for the euclidean distance $d_e$ in $\mathbb{R}^4$. 
	This function measures the distance between the endpoints irrespective of their ordering, which means that a trajectory that starts from point A and ends at point B will be close to a trajectory that starts from point B and ends at point A.
	The scale of the datasets we tested on had large enough spatial scale that clustering based on endpoints captured people moving from destination to destination, e.g. from a storefront to the sidewalk at the edge of a scene. 
	 On this dataset, distance functions that utilize the entire trajectory did not identify pedestrian intent as well as our method \cite{Morris2009,Lee2007}. It appears the metrics proposed in \cite{Morris2009} put trajectories that were similar for periods of time together even though they had different intents.
	This clustering of the end-points induces a clustering of the trajectories.
  Suppose we obtain clusters $S_1, \dots, S_n$ consisting of trajectories from our data set, as well as a set of unclassified trajectories, $S_0$.

@@

Note that the implementation of the algorithm in \cite{Kitani2012} required the endpoint of each test trajectory. 
Without this information the implementation of the predictor provided by the authors devolved into a random walk.
None of the other tested algorithms required this information.

The output distributions of the four algorithms were compared using their integrals over the cells of a regular grid over our domain. 
These integrals are used to visualize the distributions in Figures \ref{fig:bookstore-1-2} and \ref{fig:death-1-2}.
%In keeping with our measurement model, the ground truth compared against was a grid that was identically zero except on a bounding box around the measurement at time $t$. 
%The ground truth bounding box at each time instance were set equal to the average bounding box dimension centered at the 
%were the average bounding box dimensions in the training data set. 
Because our predictions all share the same scale, we amalgamate all of the prediction and truth values for all of the simulated agents at a given time step and generate ROC curves.
In our analysis, we sought a metric that evaluated the utility of prediction algorithms in the autonomous vehicle context. 
	In this instance it is critical that the generated set of predictions contains the ground-truth observed trajectory while including as few false positive detections as possible.
	ROC curves which plot the true positive rate against the false positive rate evaluate this aforementioned safety criteria for autonomous vehicles exactly. To generate the true positive rate and false positive rate, a probability distribution has to be thresholded to yield a binary map. Each chosen threshold creates a point on the ROC curve.
	The Area Under the Curve (AUC) is a standard measure of the quality of a predictor.  
	The closer that this AUC is to one, the better the prediction algorithm. We treat every value of each bounding box as a threshold.
	Figure \ref{fig:auc_vs_time} shows the analysis of the AUC of each algorithm versus time.

In addition, we evaluated the Modified Hausdorff Distance (MHD) \cite{Dubuisson1994} from the ground truth trajectory to a sample from the predictions at each time to provide a geometric measure of how accurate the predictions are. 
	Figure \ref{fig:mhd_vs_time} shows MHD plotted against time.
	Though popular in evaluating the geometric proximity of the ground truth trajectory and a predicted set, it is not the most appropriate way to evaluate the utility of an algorithm in the autonomous vehicle context. 
	Specifically consider the instance of a generated set of predictions which does not contain the ground-truth observed trajectory but is close to the ground-truth trajectory geometrically. 
	If this generated set of predictions was used in the autonomous vehicle context, it would not preclude certain portions of the space where the person was occupying. 
	Notice that the true positive rate would be zero meaning that the generated set of predictions was not useful.
	Whereas the MHD would describe this generated set of predictions as informative.
Our predictor behaves better than any of the other compared methods at moderately large $t$.
%, where we capture the behavior of the agent without suffering the misplaced confidence observed in Kitani et al.'s predictions. 
%Their encoding of velocity results in the prediction lagging behind the actual observed motion. 
This is despite providing the algorithm in \cite{Kitani2012} with the specific final location of each agent.
%Even given significantly better information, Kitani et al. generates lower quality predictions than our algorithm at all time scales.

@@

\begin{figure}
	\centering
	\includegraphics[width=0.9\linewidth]{./figures/the_results.pdf}
	\caption{A comparison of the AUC of the various algorithms. Note that the initial dip in the performance of \cite{Kitani2012} is due to their confidence in their initial estimate. We sampled the S-LSTM \cite{Alahi2016} model $100$ times to extract a less concentrated probability distribution from their algorithm.}
%	\normalmarginpar\marginnote{\#4.7}
	\label{fig:auc_vs_time}
	\vspace*{-0.75cm}
\end{figure}

\begin{figure}
	\centering
	\includegraphics[width=0.9\linewidth]{./figures/mhd_results.pdf}
	\caption{A comparison of the MHD from the ground truth of the pedestrian to a 1000 point samples from each distribution. The method from \cite{Alahi2016} does well at short time scales since it places a highly confident distribution at the given initial position of the pedestrian, but the method developed in this paper outperforms all others at intermediate times. \cite{Kitani2012}, which requires the end point of each trajectory, outperforms all other algorithms at longer time scales since they assume that the end point of the trajectory is known.}
%	\reversemarginpar\marginnote{\#4.8}
	\label{fig:mhd_vs_time}
	\vspace*{-0.7cm}
\end{figure}

@@

\section{Conclusion} 
\label{sec:conclusion}
This paper presents a real-time approach to probabilistic pedestrian modeling. 
We demonstrate the ability to accurately predict the final location of a pedestrian with superior accuracy compared to a state-of-the-art with additional contributions in two critical areas: 1) prediction speed is efficient to enable real-time control and 2) our error bound on the predictions accurately reflects potential paths for the human. 
These two features are critical in robotic applications. %
and we see the integration of such techniques with autonomous system control as one pathway to enable safe operation in human-robot spaces.
